---
title: "a3_task4_lory_salazar"
author: "Lory Salazar"
date: "11/10/2020"
output: html_document
---

```{r setup, echo = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
```

## Comparing mean size of purple urchins in control or annual sections

I will be looking at the mean sizes for the purple urchin species in Naples Reef in January 2010 to compare urchins in the control, meaning no giant kelp removal, or annual, meaning annual kelp removal, sections.

First, I read in the urchins dataset. 

```{r}
urchins <- read_csv(here("data", "urchins.csv"))
```

After reading in the dataset, I create a subset of the data to only include the purple urchins in Naples Reef observed in January 2010.

```{r}
purple_urchins <- urchins %>% 
  filter(COMMON_NAME == "Purple Urchin", SITE == "NAPL") %>% 
  mutate(date_new = mdy(DATE)) %>% 
  mutate(year = year(date_new)) %>% 
  mutate(month = month(date_new)) %>% 
  filter(month == 1, year == 2010)
```

## Looking at histograms and QQ-plots for the purple urchins dataset

After creating a subset of the data, I create histograms and QQ-plots to look at the size distribution for the two treatments: control and annual.

```{r}
ggplot(data = purple_urchins, aes(x = SIZE)) +
  geom_histogram() +
  facet_wrap(~TREATMENT) + 
  labs(x = "Size (cm)",
       y = "Purple Urchin Count")
```

Based on the two histograms, it looks like the control sample might follow a normal distribution. The annual sample is less normal. Both samples appear to vary in size of the purple urchins, without a clear pattern explaining the data. Because both samples show varied observations, there is a chance that their means are similar. 

After creating the histograms, I create qq-plots to look at the same subset of the data.

```{r}
ggplot(data = purple_urchins, aes(sample = SIZE)) +
  geom_qq() +
  facet_wrap(~TREATMENT)
```

QQ-plots are probability plots that, if showing a linear relationship, indicates a more "perfect" normal distribution Both of the qq-plots for the annual and control do not show a good linear relationship. Therefore, I would say that the annual and control groups for purple urchins do not have a normal distribution. 

## Summary Table

Next, I create a summary table of basic statistics to compare some measures of central tendency between the control and annual groups.

```{r}
purple_stats <- purple_urchins %>% 
  group_by(TREATMENT) %>% 
  summarize(mean = mean(SIZE),
            sd = sd(SIZE),
            median = median(SIZE),
            sample_size = n()
            )

purple_stats
```

From the summary table, I can see that the mean for the annual group is a bit larger than the control group. Both the standard deviation and the median of the annual group are larger as well. The sample size is almost the same for both groups. When looking back at the histograms made earlier, the summary statistics are visible; you can see that there are more large positive values for the annual group than control.

## Two-sample t-test

In this section, I perform a two-sample t-test to compare the means of the treatment groups. The test will tell us the probability of getting two sample means that are at least that different if they were drawn from populations with the same mean. The probability value will be our p-value.

```{r}
purple_control <- purple_urchins %>% 
  filter(TREATMENT == "CONTROL") %>% 
  pull(SIZE)

purple_annual <- purple_urchins %>% 
  filter(TREATMENT == "ANNUAL") %>% 
  pull(SIZE)

t.test(x = purple_control, y = purple_annual)
```

The two sample means for the annual and control groups, 3.73 and 3.92, have a mean difference of 0.190861. The data from earlier showed us that the standard deviation wasn't very different either. Because of the small difference, I would hypothesize that the samples may be coming from populations that have the same mean. After performing the t-test, I can see that the the probability value  is 25%, which is not enough evidence of differences between the population means. Because we don't have enough evidence, I would not be able to say that the samples are from populations with different means. The samples might be drawn from populations with the same means.

The p-value (0.2536) tells us that, if we draw random samples from populations with the same mean, the probability of getting two sample means that are at least as different as the ones we found here is roughly 25%.
