---
title: "a3_task3_lory_salazar"
author: "Lory Salazar"
date: "11/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
```

```{r}
urchins <- read.csv(here("data", "urchins.csv"))
```

## Section 1: Comparing purple and red urchins by size

In Section 1, I begin by exploring the urchins data and using different graphs and tables comparing the two species of urchins.

### Purple and red urchin graphs

First, I create a histogram and a box plot.

```{r}
ggplot(data = urchins, aes(x = SIZE)) +
  geom_histogram() +
  facet_wrap(~COMMON_NAME) +
  labs(x = "Size (cm)",
       y = "Urchin Count")
```

Based on these initial graphs, the red urchin sample appears to follow a fairly normal distribution. The purple urchin sample, meanwhile, is not normally distributed, and it is positively skewed, because we can see a tail in the positive direction of the x-axis. It appears that the samples might not have the same central tendencies. I would guess that the purple urchin sample has a lower measure of central tendency than red urchins.

Next, I create a box plot, which will help us to visualize the measures of central tendencies. 

```{r}
ggplot(data = urchins, aes(x = SIZE)) +
  geom_boxplot() +
  facet_wrap(~COMMON_NAME) +
  labs(x = "Size (cm)",
       y = "Count")
```

The box plots for purple and red urchins show us the differences in their samples. The purple urchin sample has a median that is more positive than that of the red urchin sample. The red urchin sample has a greater difference between the whiskers, indicating a large difference between the lowest and highest values. Additionally, both red and purple urchins have some outliers that are skewing their data positively. 

### Descriptive table

After creating graphs to explore the data, I create a summary table containing mean, median, standard deviation, and sample size. 

```{r}
urchins_stats <- urchins %>% 
  group_by(COMMON_NAME) %>% 
  summarize(mean_urchins = mean(SIZE),
            sd_urchins = sd(SIZE),
            med_urchins = median(SIZE),
            n_urchins = n())

urchins_stats
```


The summary table shows us that for all of the chosen measures of central tendency, red urchin size is larger. This aligns with the assumption I made by glancing at the histograms above. The histograms show us that the purple urchin sample is positively skewed, which is represented in the summary table here. 
The specifically, the difference in means for purple and red urchins is 2.31cm. The mean inceases by roughly 52% between the samples. 
The difference in standard deviation is smaller, only 0.72cm, while the median difference is larger, 2.5cm.

## Section 2: Exploring purple urchin counts by site and year

In section 2, I begin to explore site and year, which we previously ignored. I will be looking at the purple urchins in particular and creating both a summary table and graph to explore the data.

First, I make a subset of the data.

```{r}
purple_urchins <- urchins %>% 
  filter(COMMON_NAME == "Purple Urchin") %>% 
  mutate(date_updated = mdy(DATE)) %>% 
  mutate(year = year(date_updated)) %>% 
  mutate(site_full = case_when(
    SITE == "NAPL" ~ "Naples Reef",
    SITE == "AQUE" ~ "Arroyo Quemado Reef",
    SITE == "CARP" ~ "Carpinteria Reef",
    SITE == "MOHK" ~ "Mohawk Reef",
    SITE == "IVEE" ~ "Isla Vista Reef"
  ))
```

Next, I make a table to look at the purple urchins by site and year.

```{r}
purple_summary <- purple_urchins %>% 
  count(site_full, year)
```

Last, I create a graph to show the counts of purple urchins recorded at each site over time, from 2008-2020.

```{r}
ggplot(data = purple_summary, aes(x = year, y = n)) +
  geom_col() +
  facet_wrap(~site_full) +
  labs(x = "Time (years)",
       y = "Count of Purple Urchins")
```

